---
title: "Bibliography in R Markdown"
author: "[Phil Hurvitz](mailto:phurvitz@uw.edu)"
date: '`r format(Sys.time(), "%Y-%m-%d %H:%M")`'
header-includes: #allows you to add in your own Latex packages
- \usepackage{float} #use the 'float' package
- \floatplacement{figure}{H} #make every figure with caption = h
output: 
    bookdown::html_document2:
        number_sections: true
        self_contained: true
        code_folding: hide
        toc: true
        toc_float:
            collapsed: true
            smooth_scroll: false
    pdf_document:
        number_sections: true
        toc: true
        fig_cap: yes
        keep_tex: yes
urlcolor: blue  
bibliography: pmh.bib
csl: biomed-central.csl
link-citations: true
---

<!--these following settings control width and also make a DIV tag for black outlines for figures-->
<style type="text/css">
.main-container {
  max-width: 2400px !important;
  margin-left: auto;
  margin-right: auto;
}
</style>

<style>
.outline {
   background-color: #000000;
   padding:1px;
   display: inline-block;
}
</style>


```{r setup, message=FALSE, error=FALSE, warning=FALSE}
# key setups
library(sqldf) # to use sql syntax with data frames
library(knitr) # knitr for kable tables
library(kableExtra) # pretty tables
library(sf) # simple features (GIS)
library(leaflet) # nice maps
library(tools) # md5sum

# captions
library(captioner)
table_nums <- captioner(prefix = "Table")
figure_nums <- captioner(prefix = "Figure")

# for having DIV tags
knitr::knit_hooks$set(class = function(before, options, envir) {
  if(before){
    sprintf("<div class = '%s'>", options$class)
  }else{
    "</div>"
  }
})

knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

<!-- for a wide presentation -->
<style type="text/css">
.main-container {
  max-width: 2400px !important;
  margin-left: auto;
  margin-right: auto;
}
</style>

<!-- a DIV for black outline, good for figures-->
<style>
.outline {
   background-color: #000000;
   padding:1px;
   display: inline-block;
}
</style>

# Introduction
This example file uses the AMA-like bibliography style. For an example of APA-like bibliography, see [bibliography.html](bibliography.html) and its source code [bibliography.Rmd](bibliography.Rmd).

See the source code at the end of this document to see exactly how the references were placed in the text.

The examples of what you might see in a manuscript are highlighted in this document as __bold__ text.

Typical citations are formatted with square brackets:

```
[@Duncan2008]
```

__Some fascinating stuff was brought to light! [@Duncan2008]__

Multiple citations should be put in one set of square brackets, separated by semicolons:

```
[@Duncan2008, @Saelens2014]
```

__Two studies were completely unrelated [@Duncan2008; @Saelens2014].__

Page numbers can be included, and author can be suppressed by including a hyphen `-` before the `@` symbol (although page numbers are ignored in AMA-likce citations):

```
[-@Kang2013 pp. 1422]
```

__Kang et al. [-@Kang2013 pp. 1422] ignored PA bouts with mean accelerometry counts $\ge$ 2863 per 30 s epoch.__

In-text citations can also be used as

```
text ... @Kang2013 ... more text
```

__Kang et al. @Kang2013 ignored PA bouts with mean accelerometry counts $\ge$ 2863 per 30 s epoch.__

# Position of bibliography
The list of cited references is automatically placed at the end of the document.

For this document, because the source code should be printed after the references, we can include the following code to place the list of recerences exactly where they are wanted.

```
<div id="refs"></div>
```

# References

<div id="refs"></div>

<h2>Source code</h2>
```{r, comment='', echo=FALSE}
cat(readLines("bibliography.Rmd"), sep = '\n')
```
