---
title: "Untitled"
author: "Philip M. Hurvitz"
date: "January 21, 2021"
output: 
    bookdown::html_document2:
        number_sections: true
        self_contained: true
        code_folding: hide
        toc: true
        toc_float:
            collapsed: true
            smooth_scroll: false
    pdf_document:
        number_sections: true
        toc: true
        fig_cap: yes
        keep_tex: yes
urlcolor: blue 
---

### Equations
Equations should be numbered in manuscripts. Using `bookdown` makes this quite easy. The equations themselves require $\LaTeX$ syntax. There are numerous web sites with examples and tutorials for creating mathematical expressions with $\LaTeX$ In this example, we include Einstein's famous equation:

<pre>
\begin{equation}
  E=mc^2
  (\#eq:emc)
\end{equation}
</pre>

and the sum of squares:

<pre>
\begin{equation}
  \sum_{i=1}^n i^2 = \frac{n(n+1)(2n+1)}{6}
  (\#eq:sumn)
\end{equation}
</pre>

The label for the equation is set with `(\#eq:emc)` and can be referenced using `\@ref(eq:emc)`. Operationalized, we see:

Einstein's equation, energy equals mass times the square of the speed of light is shown in \@ref(eq:emc).

\begin{equation}
  E=mc^2
  (\#eq:emc)
\end{equation}

To make a sum of squares of _n_ first integers, see \@ref(eq:sumn).

\begin{equation}
  \sum_{i=1}^n i^2 = \frac{n(n+1)(2n+1)}{6}
  (\#eq:sumn)
\end{equation}


```{r, fig.width=3, fig.height=3}
# how many unique bins? 
bins <- length(unique(ages$age))

# create the graphic
g <- ggplot(data = ages, mapping = aes(x = age)) +
    geom_histogram(bins = bins)

# print the graphic
print(g)
```


```{r}
(A <- ages %>% 
    group_by(age) %>% 
    summarize(n = n()) %>% 
    mutate(cumsum = cumsum(n),
        "%" = round(n / sum(n) * 100, 1),
        "cum %" = round(cumsum(n / sum(n) * 100), 1)))
```


```{r}
ages %>% 
    group_by(age) %>% 
    summarize(n = n()) %>% 
    mutate(cumsum = cumsum(n),
        "%" = round(n / sum(n) * 100, 1),
        "cum %" = round(cumsum(n / sum(n) * 100), 1)) %>% 
    kable() %>% 
    kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position = "left") 
```

